---
title: "Group 3"
author: "Kaylen, Veronica Vaccaro, Lyndsey [ADD YOUR NAMES]"
date: "2025-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In plain text, someone (maybe Jadyn) should introduce our topics, data, and research questions.

```{r}
# EVERYONE LOAD YOUR LIBRARIES HERE. DO NOT DUPLICATE.
library(tidyverse)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(qdap)
library(tm)
library(SnowballC)
```

## Exploratory Analysis

Shiny App

```{r}
library(shiny)
library(ggplot2)
library(readr)
library(dplyr)

fl <- read_csv("Florida_COVID19_10302020_ByCounty_CSV_Partial.csv",
               show_col_types = FALSE) |>
  filter(COUNTYNAME != "A State") |>
  transmute(
    county = COUNTYNAME,
    NewPercPos, NewTested, NewPos,
    T_positive, T_negative, Deaths,
    MedianAge, EverMon, MonNow
  ) |>
  na.omit()

metric_choices <- c(
  "New percent positive" = "NewPercPos",
  "New tested"           = "NewTested",
  "New positives"        = "NewPos",
  "Total positives"      = "T_positive",
  "Total negatives"      = "T_negative",
  "Deaths"               = "Deaths",
  "Median age"           = "MedianAge",
  "Ever monitored"       = "EverMon",
  "Monitored now"        = "MonNow"
)

label_for <- function(var) names(metric_choices)[metric_choices == var]

ui <- fluidPage(
  titlePanel("Florida COVID-19 by County — Explorer"),
  sidebarLayout(
    sidebarPanel(
      #overview controls
      selectInput("metric", "Metric:", choices = metric_choices, selected = "NewPercPos"),
      sliderInput("topN", "Show top N counties:", min = 5, max = 25, value = 10, step = 1),
      checkboxInput("logy", "Log scale (bar & dist)", FALSE),
      tags$hr(),
      #compare (scatter) controls
      selectInput("xvar",     "X-axis:",   choices = metric_choices, selected = "MedianAge"),
      selectInput("yvar",     "Y-axis:",   choices = metric_choices, selected = "NewPercPos"),
      selectInput("sizevar",  "Point size:",  choices = metric_choices, selected = "NewTested"),
      selectInput("colorvar", "Point color:", choices = metric_choices, selected = "MedianAge"),
      checkboxInput("logx_sc", "Log X (scatter)", FALSE),
      checkboxInput("logy_sc", "Log Y (scatter)", FALSE)
    ),
    mainPanel(
      tabsetPanel(
        tabPanel("Overview (Top Counties)",
                 plotOutput("barTop", height = "360px")),
        tabPanel("Distribution",
                 plotOutput("distPlot", height = "360px")),
        tabPanel("Compare (Scatter)",
                 plotOutput("scatter", height = "420px")),
        tabPanel("Table",
                 tableOutput("tbl"))
      )
    )
  )
)

server <- function(input, output){

  output$barTop <- renderPlot({
    df <- fl |>
      arrange(desc(.data[[input$metric]])) |>
      slice_head(n = input$topN) |>
      mutate(county = reorder(county, .data[[input$metric]]))
    
    p <- ggplot(df, aes(x = county, y = .data[[input$metric]])) +
      geom_col(fill = "#2c7fb8") +
      coord_flip() +
      labs(
        title = paste0("Top ", input$topN, " counties by ", label_for(input$metric)),
        x = NULL, y = label_for(input$metric)
      ) +
      theme_minimal(base_size = 13)
    
    if (input$logy) p <- p + scale_y_log10()
    p
  })
  
  output$distPlot <- renderPlot({
    p <- ggplot(fl, aes(x = .data[[input$metric]])) +
      geom_histogram(bins = 30, alpha = 0.7) +
      geom_density(linewidth = 1) +
      labs(
        title = paste0("Distribution of ", label_for(input$metric)),
        x = label_for(input$metric), y = "Count"
      ) +
      theme_minimal(base_size = 13)
    
    if (input$logy) p <- p + scale_y_log10()
    p
  })
  
  output$scatter <- renderPlot({
    p <- ggplot(
      fl,
      aes(
        x     = .data[[input$xvar]],
        y     = .data[[input$yvar]],
        size  = .data[[input$sizevar]],
        color = .data[[input$colorvar]]
      )
    ) +
      geom_point(alpha = 0.85) +
      scale_size_continuous(range = c(2, 10)) +
      labs(
        title = "Compare two metrics by county",
        x     = label_for(input$xvar),
        y     = label_for(input$yvar),
        size  = label_for(input$sizevar),
        color = label_for(input$colorvar)
      ) +
      theme_minimal(base_size = 13)
    
    if (input$logx_sc) p <- p + scale_x_log10()
    if (input$logy_sc) p <- p + scale_y_log10()
    p
  })
  
  output$tbl <- renderTable({
    head(fl, 12)
  })
}

shinyApp(ui = ui, server = server)
```

## Linear Regression

Kingsley. Also type in plain text, explaining what you did a bit and found.

```{r}

```

## Sentiment Analysis

Lydnsey Also type in plain text, explaining what you did a bit and found.

```{r}
# Loads libraries
library(tidyverse)
library(tidytext)
library(skimr)
library(lubridate)
library(textstem)

tweets_apr_jun <- read_csv("COVIDTweetsAprilToJune2020.csv")
tweets_aug_sep <- read_csv("COVIDTweetsAprilToJune2021.csv")
tweets <- bind_rows(tweets_apr_jun, tweets_aug_sep)

# Convert created_at to date
tweets <- tweets %>%
  mutate(date = as.Date(created_at, format = "%m/%d/%Y"),
         month = floor_date(date, "month"))   # add month column

# Filters for tweets in Florida
tweets_fl <- tweets %>%
  filter(str_detect(place, regex("Florida", ignore_case = TRUE)))

# Loads Bing sentiment dictionary
bing <- get_sentiments("bing")
data("stop_words")

# Tokenize and join with Bing
joined_fl <- tweets_fl %>%
  unnest_tokens(word, clean_tweet) %>%
  mutate(word = str_to_lower(word)) %>%
  mutate(word = lemmatize_words(word)) %>%
  filter(str_detect(word, "^[a-z]+$")) %>%
  anti_join(stop_words, by = "word") %>%
  inner_join(bing, by = "word")

# Exploratory Analysis

# Quick overview
skim(tweets)

# Overall sentiment distribution
joined_fl %>%
  count(sentiment.y) %>%
  ggplot(aes(x = sentiment.y, y = n, fill = sentiment.y)) +
  geom_col() +
  labs(title = "Overall Distribution of Tweet Sentiment (Florida)")

# Monthly sentiment distribution
joined_fl %>%
  count(month, sentiment.y) %>%
  ggplot(aes(x = month, y = n, fill = sentiment.y)) +
  geom_col(position = "dodge") +
  labs(title = "Monthly Sentiment Distribution (Florida)")

# Monthly tweet volume
tweets_fl %>%
  filter(!is.na(month)) %>%
  count(month) %>%
  ggplot(aes(x = month, y = n)) +
  geom_line(color = "purple") +
  labs(title = "Monthly Tweet Volume (Florida)")

# Average monthly sentiment score
monthly_sentiment_fl <- joined_fl %>%
  mutate(sentiment_score = ifelse(sentiment.y == "positive", 1, -1)) %>%
  group_by(month) %>%
  summarise(
    avg_sentiment = mean(sentiment_score, na.rm = TRUE),
    tweet_volume = n()
  )

ggplot(monthly_sentiment_fl, aes(x = month, y = avg_sentiment)) +
  geom_line(color = "darkgreen") +
  labs(title = "Average Monthly Sentiment (Florida)")

# Top words in positive vs negative tweets
top_words_fl <- joined_fl %>%
  count(sentiment.y, word, sort = TRUE) %>%
  group_by(sentiment.y) %>%
  slice_max(n, n = 15)

ggplot(top_words_fl, aes(x = reorder(word, n), y = n, fill = sentiment.y)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment.y, scales = "free") +
  coord_flip() +
  labs(title = "Top Words in Positive vs Negative Florida Tweets")

```

## Topic Modeling

Loading my libraries and creating custom stop words.

```{r}

more_stop_words <- tribble(
  ~word, ~lexicon,
  "covid19", "CUSTOM",
  "coronavirus", "CUSTOM",
  "rt", "CUSTOM",
  "covid", "CUSTOM",
  "19", "CUSTOM",
  "florida", "CUSTOM",
  "coronaviru", "CUSTOM",
  "http", "CUSTOM",
  "covidnineteen", "CUSTOM",
  "don’t", "CUSTOM",
  "it’s", "CUSTOM",
  "hundreds", "CUSTOM",
  "thousands", "CUSTOM",
  "millions", "CUSTOM",
  "thousandth", "CUSTOM",
  "fauci", "CUSTOM",
  "amp", "CUSTOM",
  "port", "CUSTOM"
)

full_stop_words <- stop_words %>% 
  bind_rows(more_stop_words)

```

April to June of 2020 - Preparing my data

```{r}

  data2020 <- read.csv("COVIDTweetsAprilToJune2020.csv")
  
  original_2020 <- data2020 %>% 
    filter(str_detect(place, "Florida|FL")) %>% 
    select(id, original_text)
  
  # cleaning 2020 original tweets FROM SCRATCH
  clean2020 <- original_2020 %>% 
    mutate(original_text = as.character(original_text)) %>% 
    mutate(original_text = replace_contraction(original_text)) %>% 
    mutate(original_text = replace_number(original_text)) %>% 
    mutate(original_text = str_remove_all(original_text, "https?://.*$")) %>% 
    mutate(original_text = str_remove_all(original_text, "RT ")) %>% 
    mutate(original_text = str_remove_all(original_text, "@\\w+")) %>% 
    mutate(original_text = removePunctuation(original_text)) %>%
    mutate(original_text = str_remove_all(original_text, "'")) %>% 
    mutate(original_text = removeNumbers(original_text)) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words)

  dictionary_2020 <- original_2020 %>% 
    select(original_text) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words) %>%
    distinct(word) %>% 
    pull(word)
  # This dictionary will be used to stem-complete the words after I stem them.
  
  stem2020 <- clean2020 %>% 
    mutate(word = wordStem(word)) %>% 
    mutate(
      word = unlist(lapply(word, stemCompletion, dictionary = dictionary_2020))
    )
  
  # The words did not 100% stem the way I had hoped, so I manually fixed some of them after searching/filtering through the original tweets to find the best match if needed. I repeat these same words for manual stem completion in the 2021 data processing (next) too. Still, stem completion saved a lot of time and cleaned up the data well.
  words_2020 <- stem2020 %>% 
    drop_na() %>% 
    anti_join(full_stop_words) %>% 
    mutate(word = recode(word,
                         "viru" = "virus",
                         "informat" = "information",
                         "posit" = "positive",
                         "mani" = "humanity",
                         "commun" = "community",
                         "preside" = "president", 
                         "respons" = "response",
                         "raise" = "fundraise",
                         "busin" = "business",
                         "hospita" = "hospital",
                         "vaccin" = "vaccine",
                         "accounc" = "announce",
                         "countie" = "county",
                         "populati" = "population",
                         "emerge" = "emergency",
                         "announc" = "announce")) %>% 
    count(id, word) %>% 
    filter(n > 10) # There were a lot of words from this tweet dataset, so I filtered for only the words that appeared 10 or more times. In the 2021 data, this part is very different.

```

Creating the 2020 LDA model:

```{r}

  dtm2020 <- words_2020 %>% 
    cast_dtm(id, word, n)
  
  lda2020_k16 = LDA(
    dtm2020,
    k = 16, # 16 was the most optimal, human-interpretable k value, according to me (Kaylen) and Jadyn, for this model.
    method = "Gibbs",
    control = list(seed = 67)
  )
  
  beta_2020_k16 <- lda2020_k16 %>% 
    tidy(matrix = "beta") # Beta to find out the words within a topic to analyze them.
  
  beta_2020_k16 <- beta_2020_k16 %>% 
    arrange(desc(beta)) %>% 
    group_by(topic) %>%
    slice_max(beta, n = 8) %>% # 8 words at a time in a column chart, but if any beta values tie, they will both (or more) show, which is fine.
    mutate(term2 = fct_reorder(term, beta))
  
  ggplot(beta_2020_k16, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") +
    coord_flip() +
    labs(title = "April - June 2020 Topics",
         x = "Words",
         y = "Beta")
  
  # Labeling human-identified topics using k = 16
  beta_2020_k16 <- beta_2020_k16 %>% 
    filter(topic %in% 1:12) %>% 
    mutate(
      topic = as.character(topic),
      topic = recode(topic, # After visualizing with the ggplot code above, Jadyn and I came up with these topics.
                     "1" = "American Healthcare",
                     "2" = "Government Response",
                     "3" = "Mortality & Global Severity",
                     "4" = "Daily Reports on Covid Numbers",
                     "5" = "Community Support",
                     "6" = "Mask Usage",
                     "7" = "Public Media Criticism",
                     "8" = "Vaccine Advancement Progress",
                     "9" = "Economic Crisis",
                     "10" = "Trump and Stimulus Checks",
                     "11" = "CDC Virus Information",
                     "12" = "Quarantine",
                     "13" = "Global Situation",
                     "14" = "Reopening the Workplace",
                     "15" = "Reopening Impact",
                     "16" = "Strain on Hospitals & Healthcare"))
  
  # Replot with 16 topic names
  ggplot(beta_2020_k16, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + # free_y scale so that the beta ticks change in the same increments across plots.
    coord_flip() +
    labs(title = "April - June 2020 Topics",
         x = "Words",
         y = "Beta")
  
  gamma_2020_k16 <- lda2020_k16 %>% 
    tidy(matrix = "gamma") %>% # Gamma to find out which topics were most dominant.
    mutate(
      topic = as.character(topic),
      topic = recode(topic,
                     "1" = "American Healthcare",
                     "2" = "Government Response",
                     "3" = "Mortality & Global Severity",
                     "4" = "Daily Reports on Covid Numbers",
                     "5" = "Community Support",
                     "6" = "Mask Usage",
                     "7" = "Public Media Criticism",
                     "8" = "Vaccine Advancement Progress",
                     "9" = "Economic Crisis",
                     "10" = "Trump and Stimulus Checks",
                     "11" = "CDC Virus Information",
                     "12" = "Quarantine",
                     "13" = "Global Situation",
                     "14" = "Reopening the Workplace",
                     "15" = "Reopening Impact",
                     "16" = "Strain on Hospitals & Healthcare"))
  
  gamma_2020_k16 <- gamma_2020_k16 %>% 
    group_by(topic) %>% 
    summarize(gamma_total = sum(gamma)) %>% 
    arrange(desc(gamma_total))
  
  gamma_2020_k16
  
  ggplot(gamma_2020_k16, aes(x="", y=gamma_total, fill=topic)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar("y", start=0) +
    theme_void() +
    labs(
      title = "Most Popular Topic: Daily Reports On Covid Numbers",
      subtitle = "April-June 2020",
      fill = "Topics"
    )

```
The most prominent topics that Floridians, likely like most people, talked about during April-June of 2020 are:
1. **Daily Reports on Covid Numbers** - likely because cases keep stacking and it brews fear
2. **Mortality & Global Severity** - a global pandemic was definitely a new experience for most people alive today. The coronavirus caused severe illness and many deaths. It was happening globally and hitting close to many people's homes and personal lives. 
3. **Quarantine** - At this point in time, Florida had just begun reopening its counties, allowing public places such as retail stores and restuarants. 


I basically repeat most of the code I did for the 2020 data with the 2021 data.
April to June of 2021 - Preparing my data:

```{r}

  data2021 <- read.csv("COVIDTweetsAprilToJune2021.csv")
  
  original_2021 <- data2021 %>% 
    filter(str_detect(place, "Florida|FL")) %>% 
    select(id, original_text)
  
  # cleaning 2021 original tweets FROM SCRATCH
  clean2021 <- original_2021 %>% 
    mutate(original_text = as.character(original_text)) %>% 
    mutate(original_text = replace_contraction(original_text)) %>% 
    mutate(original_text = replace_number(original_text)) %>% 
    mutate(original_text = str_remove_all(original_text, "https?://.*$")) %>% 
    mutate(original_text = str_remove_all(original_text, "RT ")) %>% 
    mutate(original_text = str_remove_all(original_text, "@\\w+")) %>% 
    mutate(original_text = removePunctuation(original_text)) %>%
    mutate(original_text = str_remove_all(original_text, "'")) %>% 
    mutate(original_text = removeNumbers(original_text)) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words)
  
  dictionary_2021 <- original_2021 %>% 
    select(original_text) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words) %>%
    distinct(word) %>% 
    pull(word)
  
  stem2021 <- clean2021 %>% 
    mutate(word = wordStem(word)) %>% 
    mutate(
      word = unlist(lapply(word, stemCompletion, dictionary = dictionary_2021))
    )
  
  words_2021 <- stem2021 %>% 
    drop_na() %>% 
    anti_join(full_stop_words) %>% 
    mutate(word = recode(word,
                         "viru" = "virus",
                         "informat" = "information",
                         "posit" = "positive",
                         "mani" = "humanity",
                         "commun" = "community",
                         "preside" = "president", 
                         "respons" = "response",
                         "raise" = "fundraise",
                         "busin" = "business",
                         "hospita" = "hospital",
                         "vaccin" = "vaccine",
                         "accounc" = "announce",
                         "countie" = "county",
                         "populati" = "population",
                         "emerge" = "emergency",
                         "announc" = "announce")) %>% 
    count(id, word) %>% 
    filter(n > 2)

```

Creating the 2021 LDA model:

```{r}

  dtm2021 <- words_2021 %>% 
    cast_dtm(id, word, n)
  
  lda2021_k10 = LDA(
    dtm2021,
    k = 10, # 10 was the most optimal, human-interpretable k value, according to me (Kaylen) and Jadyn, for this model
    method = "Gibbs",
    control = list(seed = 67)
  )
  
  beta_2021_k10 <- lda2021_k10 %>% 
    tidy(matrix = "beta")
  
  beta_2021_k10 <- beta_2021_k10 %>% 
    arrange(desc(beta)) %>% 
    group_by(topic) %>%
    filter(beta >= .0125) %>% 
    slice_max(beta, n = 8) %>% 
    mutate(term2 = fct_reorder(term, beta))
  
  ggplot(beta_2021_k10, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + 
    coord_flip() +
    labs(
      title = "April - June 2021 Topics",
      x = "Words",
      y = "Beta"
    )
  
  # Labeling human-identified topics using k = 10
  beta_2021_k10 <- beta_2021_k10 %>% 
    filter(topic %in% 1:12) %>% 
    mutate(
      topic = as.character(topic),
      topic = recode(topic,
                     "1" = "Vaccine Announcements & Expert Advice",
                     "2" = "Vaccine Distribution & County-level Reporting",
                     "3" = "Negative Effects of the Pandemic",
                     "4" = "Vaccine Transmission Messageing",
                     "5" = "Market & Economy",
                     "6" = "Vaccine & Public Health Measures",
                     "7" = "Political Framing of Vaccines",
                     "8" = "Vaccine Approval",
                     "9" = "Health Reports",
                     "10" = "State of Covid Vaccines in June"))
  
  # Replot with 10 topic names
  ggplot(beta_2021_k10, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + 
    coord_flip() +
    labs(
      title = "April - June 2021 Topics",
      x = "Words",
      y = "Beta"
    )
  
  gamma_2021_k10 <- lda2021_k10 %>% 
    tidy(matrix = "gamma") %>% 
    mutate(
      topic = as.character(topic),
      topic = recode(topic,
                     "1" = "Vaccines",
                     "2" = "Vaccines",
                     "3" = "Negative Effects of the Pandemic",
                     "4" = "Vaccines",
                     "5" = "Market & Economy",
                     "6" = "Vaccines",
                     "7" = "Vaccines",
                     "8" = "Vaccines",
                     "9" = "Health Reports",
                     "10" = "Vaccines"))
  
  gamma_2021_k10 <- gamma_2021_k10 %>% 
    group_by(topic) %>% 
    summarize(gamma_total = sum(gamma)) %>% 
    arrange(desc(gamma_total))
  
  gamma_2021_k10

  ggplot(gamma_2021_k10, aes(x="", y=gamma_total, fill=topic)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar("y", start=0) +
    theme_void() +
    labs(
      title = "Most Popular Topic: Covid Vaccinations",
      subtitle = "April-June 2021",
      fill = "Topics"
    )
```
Though we set our k value to equal ten and ended up labeling ten topics, one thing I noticed about the 2021 tweet data was that the word "vaccine" dominated in several of the topics when observing beta values. It's clear that the conversation shifted to the topics of covid vaccinations in 2021. This contrasts the conversations around covid in 2020, where we were living in a time of __starting__ to return to normal routines after a strict quarantine, lived through reports of many cases, thinking the world changed forever. Conversations were more diverse around the main topic of Covid back then. In 2021, when vaccines were out and rolling, Governor Desantis vocally spoke about no vaccine mandates.

## Conclusion

Jadyn.
