---
title: "Group 3"
author: "Kaylen, Veronica Vaccaro, Lyndsey [ADD YOUR NAMES]"
date: "2025-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In plain text, someone (maybe Jadyn) should introduce our topics, data, and research questions.

```{r}
# EVERYONE LOAD YOUR LIBRARIES HERE. DO NOT DUPLICATE.
library(tidyverse)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(qdap)
library(tm)
library(SnowballC)
library(shiny)
library(readr)
library(skimr)
library(lubridate)
library(textstem)
```

## Exploratory Analysis

For this exploratory analysis, I used a Shiny app to visualoze COVID-19 data accross Florida Counties. I selected the most relevant variables such as cases, deaths, testing, monitoring, and median age. I also removed missing values to make sure there would be no erros. Through the usage of Shiny app, I was able to compare metrics interactively through bar charts, histograms, and scatter plots. I was able to find that larger urban counties had the highest total cases, but small rural counties had higher positivity rates, suggesting that county density and testing capacity play a role in detecting cases. Scatterplots comparing median age and deaths showed that older, more populous counties showed mortality. 

```{r}

fl <- read_csv(
  "Florida_COVID19_10302020_ByCounty_CSV_Partial.csv",
  show_col_types = FALSE
) |>
  filter(COUNTYNAME != "A State") |>
  transmute(
    county = COUNTYNAME,      # nicer name for UI
    NewPercPos,               # new percent positive 
    NewTested,                # count new tests
    NewPos,                   # count new positives
    T_positive,               # cumulative positives
    T_negative,               # cumulative negatives
    Deaths,                   # cumulative deaths
    MedianAge,                # county median age
    EverMon,                  # ever monitored
    MonNow                    # monitored now
  ) |>
  na.omit()

#Shared labels for UI controls 
#Named vector: user-facing labels on the left, column names on the right.
metric_choices <- c(
  "New percent positive" = "NewPercPos",
  "New tested"           = "NewTested",
  "New positives"        = "NewPos",
  "Total positives"      = "T_positive",
  "Total negatives"      = "T_negative",
  "Deaths"               = "Deaths",
  "Median age"           = "MedianAge",
  "Ever monitored"       = "EverMon",
  "Monitored now"        = "MonNow"
)

#Small helper to convert a selected column name back to its label
label_for <- function(var) names(metric_choices)[metric_choices == var]

#UI 
ui <- fluidPage(
  titlePanel("Florida COVID-19 by County — Explorer"),
  sidebarLayout(
    sidebarPanel(
      #Overview (bar + dist) controls
      selectInput("metric", "Metric:", choices = metric_choices, selected = "NewPercPos"),
      sliderInput("topN", "Show top N counties:", min = 5, max = 25, value = 10, step = 1),
      tags$hr(),
      #Compare (scatter) controls
      selectInput("xvar",     "X-axis:",     choices = metric_choices, selected = "MedianAge"),
      selectInput("yvar",     "Y-axis:",     choices = metric_choices, selected = "NewPercPos"),
      selectInput("sizevar",  "Point size:", choices = metric_choices, selected = "NewTested"),
      selectInput("colorvar", "Point color:",choices = metric_choices, selected = "MedianAge")
    ),
    mainPanel(
      tabsetPanel(
        tabPanel("Overview (Top Counties)", plotOutput("barTop", height = "360px")),
        tabPanel("Distribution",            plotOutput("distPlot", height = "360px")),
        tabPanel("Compare (Scatter)",       plotOutput("scatter", height = "420px")),
        tabPanel("Table",                   tableOutput("tbl"))
      )
    )
  )
)

#Server
server <- function(input, output){
  
  #Top-N bar chart of the selected metric
  output$barTop <- renderPlot({
    #Rank counties by selected metric and keep the top N
    df <- fl |>
      arrange(desc(.data[[input$metric]])) |>
      slice_head(n = input$topN) |>
      mutate(county = reorder(county, .data[[input$metric]]))  #order bars
    
    p <- ggplot(df, aes(x = county, y = .data[[input$metric]])) +
      geom_col(fill = "#2c7fb8") +
      coord_flip() +
      labs(
        title = paste0("Top ", input$topN, " counties by ", label_for(input$metric)),
        x = NULL,
        y = label_for(input$metric)
      ) +
      theme_minimal(base_size = 13)
    
    p
  })
  
  #Distribution (histogram + density) of the selected metric
  output$distPlot <- renderPlot({
    p <- ggplot(fl, aes(x = .data[[input$metric]])) +
      geom_histogram(bins = 30, alpha = 0.7) +     # histogram for frequency
      geom_density(linewidth = 1) +                # smoothed shape overlay
      labs(
        title = paste0("Distribution of ", label_for(input$metric)),
        x = label_for(input$metric),
        y = "Count"
      ) +
      theme_minimal(base_size = 13)
    p
  })
  
  #Scatter: compare any two metrics, with size/color encodings 
  output$scatter <- renderPlot({
    p <- ggplot(
      fl,
      aes(
        x     = .data[[input$xvar]],
        y     = .data[[input$yvar]],
        size  = .data[[input$sizevar]],   # third variable via point size
        color = .data[[input$colorvar]]   # fourth variable via color
      )
    ) +
      geom_point(alpha = 0.85) +
      scale_size_continuous(range = c(2, 10)) +
      labs(
        title = "Compare two metrics by county",
        x     = label_for(input$xvar),
        y     = label_for(input$yvar),
        size  = label_for(input$sizevar),
        color = label_for(input$colorvar)
      ) +
      theme_minimal(base_size = 13)
    p
  })
  
  #Preview table
  output$tbl <- renderTable({
    head(fl, 12)
  })
}

#Run app 
shinyApp(ui = ui, server = server)
```

## Linear Regression

For this regression Model, I selected the key variables, removed the missing values, and built a predicting model that analyzed which county factors were related to new percent positive cases. The results show that Median Age had the strongest relationship with New Percent Positive. At the same time, the other variables (EverMon and MonNow) did not predict changes in positivity rates as well as Median Age.

```{r}
# Loaded packages are above


# Read the dataset 
myDF <- read.csv("Florida_COVID19_10302020_ByCounty_CSV_Partial.csv")



# Keep relevant variables
myDF2 <- myDF[, c("COUNTYNAME",
                  "NewPercPos",
                  "MedianAge",
                  "EverMon",
                  "MonNow")]

# Remove missing values
myDF2 <- na.omit(myDF2)

# Scatterplot of New Percent Positive vs Median Age
plot(myDF2$MedianAge, myDF2$NewPercPos,
     xlab = "Median Age",
     ylab = "New Percent Positive",
     main = "New Percent Positive vs Median Age")

# Linear regression model
lm1 <- lm(NewPercPos ~ MedianAge + EverMon + MonNow, data = myDF2)

# Model summary
summary(lm1)

# Simpler model
lm2 <- lm(NewPercPos ~ MedianAge, data = myDF2)
summary(lm2)

# Fitted vs Observed plot
plot(fitted(lm1), myDF2$NewPercPos,
     xlab = "Fitted Values",
     ylab = "Observed New Percent Positive",
     main = "Fitted vs Observed")
abline(0, 1)
```

## Sentiment Analysis

Lydnsey Also type in plain text, explaining what you did a bit and found.

```{r}
# Loads libraries
tweets_apr_jun <- read_csv("COVIDTweetsAprilToJune2020.csv")
tweets_aug_sep <- read_csv("COVIDTweetsAprilToJune2021.csv")
tweets <- bind_rows(tweets_apr_jun, tweets_aug_sep)

# Convert created_at to date
tweets <- tweets %>%
  mutate(date = as.Date(created_at, format = "%m/%d/%Y"),
         month = floor_date(date, "month"))   # add month column

# Filters for tweets in Florida
tweets_fl <- tweets %>%
  filter(str_detect(place, regex("Florida", ignore_case = TRUE)))

# Loads Bing sentiment dictionary
bing <- get_sentiments("bing")
data("stop_words")

# Tokenize and join with Bing
joined_fl <- tweets_fl %>%
  unnest_tokens(word, clean_tweet) %>%
  mutate(word = str_to_lower(word)) %>%
  mutate(word = lemmatize_words(word)) %>%
  filter(str_detect(word, "^[a-z]+$")) %>%
  anti_join(stop_words, by = "word") %>%
  inner_join(bing, by = "word")

# Exploratory Analysis

# Quick overview
skim(tweets)

# Overall sentiment distribution
joined_fl %>%
  count(sentiment.y) %>%
  ggplot(aes(x = sentiment.y, y = n, fill = sentiment.y)) +
  geom_col() +
  labs(title = "Overall Distribution of Tweet Sentiment (Florida)")

# Monthly sentiment distribution
joined_fl %>%
  count(month, sentiment.y) %>%
  ggplot(aes(x = month, y = n, fill = sentiment.y)) +
  geom_col(position = "dodge") +
  labs(title = "Monthly Sentiment Distribution (Florida)")

# Monthly tweet volume
tweets_fl %>%
  filter(!is.na(month)) %>%
  count(month) %>%
  ggplot(aes(x = month, y = n)) +
  geom_line(color = "purple") +
  labs(title = "Monthly Tweet Volume (Florida)")

# Average monthly sentiment score
monthly_sentiment_fl <- joined_fl %>%
  mutate(sentiment_score = ifelse(sentiment.y == "positive", 1, -1)) %>%
  group_by(month) %>%
  summarise(
    avg_sentiment = mean(sentiment_score, na.rm = TRUE),
    tweet_volume = n()
  )

ggplot(monthly_sentiment_fl, aes(x = month, y = avg_sentiment)) +
  geom_line(color = "darkgreen") +
  labs(title = "Average Monthly Sentiment (Florida)")

# Top words in positive vs negative tweets
top_words_fl <- joined_fl %>%
  count(sentiment.y, word, sort = TRUE) %>%
  group_by(sentiment.y) %>%
  slice_max(n, n = 15)

ggplot(top_words_fl, aes(x = reorder(word, n), y = n, fill = sentiment.y)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment.y, scales = "free") +
  coord_flip() +
  labs(title = "Top Words in Positive vs Negative Florida Tweets")

```

## Topic Modeling

The goal is to compare topics and themes discussed revolving around COVID-19 during April-June of 2020 and April-June of 2021---a one year difference.

**I created my own customs words.** Some of the words are obvious topic words, such as "covid19" which would become redundant. Others include numbers that didn't get cleaned out with my cleaning functions and some other words that add no value to the topic-naming stage via beta scores.

```{r}

more_stop_words <- tribble(
  ~word, ~lexicon,
  "covid19", "CUSTOM",
  "coronavirus", "CUSTOM",
  "rt", "CUSTOM",
  "covid", "CUSTOM",
  "19", "CUSTOM",
  "florida", "CUSTOM",
  "coronaviru", "CUSTOM",
  "http", "CUSTOM",
  "covidnineteen", "CUSTOM",
  "don’t", "CUSTOM",
  "it’s", "CUSTOM",
  "hundreds", "CUSTOM",
  "thousands", "CUSTOM",
  "millions", "CUSTOM",
  "thousandth", "CUSTOM",
  "fauci", "CUSTOM",
  "amp", "CUSTOM",
  "port", "CUSTOM",
  "due", "CUSTOM",
  "stainless", "CUSTOM",
  "include", "CUSTOM",
  "video", "CUSTOM",
  "dyannleroy", "CUTSOM",
  "ten", "CUSTOM",
  "million", "CUSTOM",
  "di", "CUSTOM",
  "pm", "CUSTOM",
  "dianeswonk", "CUSTOM",
  "residency", "CUSTOM"
)

full_stop_words <- stop_words %>% 
  bind_rows(more_stop_words)

```

### April to June of 2020 - Preparing my data
- I loaded in my data, assigned unique ids, cleaned the tweets, tokenized my documents, performed word stemming and stem completion, the counted words per id.

```{r}

  data2020 <- read.csv("COVIDTweetsAprilToJune2020.csv")
  
  original_2020 <- data2020 %>% 
    filter(str_detect(place, "Florida|FL")) %>% 
    select(id, original_text) %>% 
    mutate(id = row_number()) # Reassign unique ids because the scientific notation made documents appears to be the same docs/source.
  
  # cleaning 2020 original tweets FROM SCRATCH
  clean2020 <- original_2020 %>% 
    mutate(original_text = as.character(original_text)) %>% 
    mutate(original_text = replace_contraction(original_text)) %>% 
    mutate(original_text = replace_number(original_text)) %>% 
    mutate(original_text = str_remove_all(original_text, "https?://.*$")) %>% 
    mutate(original_text = str_remove_all(original_text, "RT ")) %>% 
    mutate(original_text = str_remove_all(original_text, "@\\w+")) %>% 
    mutate(original_text = removePunctuation(original_text)) %>%
    mutate(original_text = str_remove_all(original_text, "'")) %>% 
    mutate(original_text = removeNumbers(original_text)) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words)

  dictionary_2020 <- original_2020 %>% 
    select(original_text) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words) %>%
    distinct(word) %>% 
    pull(word)
  # This dictionary will be used to stem-complete the words after I stem them.
  
  stem2020 <- clean2020 %>% 
    mutate(word = wordStem(word)) %>% 
    mutate(
      word = unlist(lapply(word, stemCompletion, dictionary = dictionary_2020))
    )
  # The words did not 100% stem the way I had hoped, so I manually fixed some of them after searching/filtering through the original tweets to find the best match if needed. I repeat these same words for manual stem completion in the 2021 data processing (next) too. Still, stem completion saved a lot of time and cleaned up the data well enough.
  words_2020 <- stem2020 %>% 
    drop_na() %>% 
    anti_join(full_stop_words) %>% 
    mutate(word = recode(word,
                         "viru" = "virus",
                         "informat" = "information",
                         "posit" = "positive",
                         "mani" = "humanity",
                         "commun" = "community",
                         "preside" = "president", 
                         "respons" = "response",
                         "raise" = "fundraise",
                         "busin" = "business",
                         "hospita" = "hospital",
                         "vaccin" = "vaccine",
                         "accounc" = "announce",
                         "countie" = "county",
                         "populati" = "population",
                         "emerge" = "emergency",
                         "announc" = "announce",
                         "updat" = "update")) %>% 
    count(id, word)

```

### Creating the 2020 LDA model:

```{r}

  # LDA Modeling
  dtm2020 <- words_2020 %>% 
    cast_dtm(id, word, n)
  
  lda2020_k9 = LDA(
    dtm2020,
    k = 9, # After many trial and errors, I found this k value to be the most humanly interpret-able for the 2020 covid data.
    method = "Gibbs",
    control = list(seed = 67)
  )
  
  glimpse(lda2020_k9)
  
  beta_2020_k9 <- lda2020_k9 %>% 
    tidy(matrix = "beta") # Beta to find out the words within a topic to analyze them.
  
  beta_2020_k9 <- beta_2020_k9 %>% 
    arrange(desc(beta)) %>% 
    group_by(topic) %>%
    slice_max(beta, n = 8) %>% # 8 words at a time in a column chart, but if any beta values tie, they will both (or more) show, which is fine.
    mutate(term2 = fct_reorder(term, beta))
  
  ggplot(beta_2020_k9, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + # free_y scale so that the beta ticks change in the same increments across plots.
    coord_flip() +
    labs(title = "April - June 2020 Topics",
         x = "Words",
         y = "Beta")
  
  # Labeling human-identified topics using k = 9
  beta_2020_k9 <- beta_2020_k9 %>% 
    mutate(
      topic = as.character(topic),
      topic = recode(topic,
                     "1" = "Severity Reports",
                     "2" = "Trump Response",
                     "3" = "National Government Response",
                     "4" = "Global Supplies",
                     "5" = "Vaccine Development",
                     "6" = "Early Reopening Effects",
                     "7" = "Government Action",
                     "8" = "Virus Exposure to Workers",
                     "9" = "Closing of American Public Spaces"))
  
  # Re-plot w/ topic titles
  ggplot(beta_2020_k9, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + 
    coord_flip() +
    labs(title = "April - June 2020 Topics",
         x = NULL,
         y = "Beta") + theme_bw()
  
  # gamma to find most dominant/prevalent topic during 2020
  gamma_2020_k16 <- lda2020_k9 %>% 
    tidy(matrix = "gamma") %>%
    mutate(
      topic = recode(topic,
                     "1" = "Severity Reports",
                     "2" = "Trump Response",
                     "3" = "National Government Response",
                     "4" = "Global Supplies",
                     "5" = "Vaccine Development",
                     "6" = "Early Reopening Effects",
                     "7" = "Government Action",
                     "8" = "Virus Exposure to Workers",
                     "9" = "Closing of American Public Spaces"))
  
  dominant_topic_2020 <- gamma_2020_k16 %>%
    group_by(document) %>%
    slice_max(gamma, n = 1, with_ties = FALSE) %>%
    ungroup() # finding the best fit topic per document
  
  topic_counts_2020 <- dominant_topic_2020 %>% 
    count(topic, sort = TRUE) %>% 
    mutate(topic = reorder(topic, n)) # adding up documents per topic
  
  # Bar charts > Pie charts
  ggplot(topic_counts_2020, aes(x = topic, y = n, fill = topic)) +
    geom_bar(stat = "identity", width = 1, color = "white", show.legend = F) +
    geom_text(aes(label = n)) +
    coord_flip() +
    labs(
      title = "The Biggest Themes of April - June 2020",
      subtitle = "Relating to Covid-19",
      fill = "Themes",
      x = NULL,
      y = "Tweets"
    ) + theme_bw()

```
**For context**: Early April, the CDC announced new mask-wearing guidelines. Many people complied; current President Trump did not. Florida ended quarantine on April 30th and moved into phases of reopening the public. In early June, the state had entered phase two of reopening, meaning allowing bars, theaters, and other public building spaces to run at 50% capacity. There was no vaccine yet, but development, research, and human clinical trials had been ongoing and fast. 

The most prominent topics that Floridians, likely like most people, talked about during April-June of 2020 are:
1. **Severity Reports** - Because the global pandemic is new to most people, and many people across the globe were experiencing severe illness and death counts kept rising, there were so many updates. I remember back in my senior of high school where I was checking covid cases every day. With just one case in your state, it spreads like wildfire---both the virus and news.
2. **Trump's Response** - We were under Trump's presidency during the time these tweets were collected. Many people criticized his response to covid-19, where he would give poor medical advice (such as injecting disinfectants to treat the virus at home), him turning the U.S. away from participating in the World Health Organization's vaccine project, and his lack of serious leadership.
3. **National Government Response** - During this period, the White House held frequent briefings and released evolving public health guidelines. Additionally, during March, Medicaid temporarily expanded telehealth benefits. This provided a wider range of healthcare to the beneficiaries with the need to travel to a healthcare facility.

***More can be found here:*** <https://www.cdc.gov/museum/timeline/covid19.html>

Overall, many of the tweets consisted of reporting of official news, health guidelines and also a lot of conversations around these new experiences, such as the entire pandemic and also reopening after quarantine, which some people were precautious and others were tired of quarantine and masks.

One topic that I found interesting and nostalgic (but I changed the model since I made a *major* preprocessing error) was something Jadyn and I labeled as *Community Support for Healthcare Workers*. After code revisions and changes in the k value, that topic was never seen or at least obvious again.

### April to June of 2021 - Preparing my data and running an LDA model
I basically repeat most of the code I did for the 2020 data with the 2021 data, switching out and renaming new objects.

```{r}

 data2021 <- read.csv("COVIDTweetsAprilToJune2021.csv")
  
  original_2021 <- data2021 %>% 
    filter(str_detect(place, "Florida|FL")) %>% 
    select(id, original_text) %>% 
    mutate(id = row_number())
  
  # cleaning 2021 original tweets FROM SCRATCH
  clean2021 <- original_2021 %>% 
    mutate(original_text = as.character(original_text)) %>% 
    mutate(original_text = replace_contraction(original_text)) %>% 
    mutate(original_text = replace_number(original_text)) %>% 
    mutate(original_text = str_remove_all(original_text, "https?://.*$")) %>% 
    mutate(original_text = str_remove_all(original_text, "RT ")) %>% 
    mutate(original_text = str_remove_all(original_text, "@\\w+")) %>% 
    mutate(original_text = removePunctuation(original_text)) %>%
    mutate(original_text = str_remove_all(original_text, "'")) %>% 
    mutate(original_text = removeNumbers(original_text)) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words)
  
  dictionary_2021 <- original_2021 %>% 
    select(original_text) %>% 
    unnest_tokens(word, original_text) %>% 
    anti_join(full_stop_words) %>%
    distinct(word) %>% 
    pull(word)
  
  stem2021 <- clean2021 %>% 
    mutate(word = wordStem(word)) %>% 
    mutate(
      word = unlist(lapply(word, stemCompletion, dictionary = dictionary_2021))
    )
  
  words_2021 <- stem2021 %>% 
    drop_na() %>% 
    anti_join(full_stop_words) %>% 
    mutate(word = recode(word,
                         "viru" = "virus",
                         "informat" = "information",
                         "posit" = "positive",
                         "mani" = "humanity",
                         "communi" = "community",
                         "preside" = "president", 
                         "respons" = "response",
                         "raise" = "fundraise",
                         "busin" = "business",
                         "hospita" = "hospital",
                         "vaccin" = "vaccine",
                         "accounc" = "announce",
                         "countie" = "county",
                         "populati" = "population",
                         "emerge" = "emergency",
                         "announc" = "announce",
                         "updat" = "update",
                         "requir" = "require")) %>% 
    count(id, word)
  
  # LDA Modeling
  dtm2021 <- words_2021 %>% 
    cast_dtm(id, word, n)
  
  lda2021_k10 = LDA(
    dtm2021,
    k = 8, 
    method = "Gibbs",
    control = list(seed = 10)
  )
  
  glimpse(lda2021_k10)
  
  beta_2021_k10 <- lda2021_k10 %>% 
    tidy(matrix = "beta")
  
  beta_2021_k10 <- beta_2021_k10 %>% 
    arrange(desc(beta)) %>% 
    group_by(topic) %>%
    slice_max(beta, n = 8) %>% 
    mutate(term2 = fct_reorder(term, beta))
  
  ggplot(beta_2021_k10, aes(term, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + 
    coord_flip() +
    labs(
      title = "April - June 2021 Topics",
      x = "Words",
      y = "Beta"
    )
  
  # Labeling human-identified topics using k = 10
  beta_2021_k10 <- beta_2021_k10 %>% 
    mutate(
      topic = as.character(topic),
      topic = recode(topic,
                     "1" = "Daily Vaccine Updates",
                     "2" = "Global Vaccine Rollout",
                     "3" = "Variants and Lab Findings",
                     "4" = "Research Updates",
                     "5" = "Administrating Vaccines",
                     "6" = "Student Vaccination",
                     "7" = "National Covid Death Reports",
                     "8" = "Vaccine Approval"))
  
  # # Re-plot w/ topic titles
  ggplot(beta_2021_k10, aes(term2, beta, fill = as.factor(topic))) + 
    geom_col(show.legend = F) +
    facet_wrap(~topic, scales = "free_y") + 
    coord_flip() +
    labs(
      title = "April - June 2021 Topics",
      x = NULL,
      y = "Beta"
    ) + theme_bw()
  
  # gamma to find most dominant/prevalent topic during 2020
  gamma_2021_k10 <- lda2021_k10 %>% 
    tidy(matrix = "gamma") %>% 
    mutate(
    topic = recode(topic,
                   "1" = "Daily Vaccine Updates",
                   "2" = "Global Vaccine Rollout",
                   "3" = "Variants and Lab Findings",
                   "4" = "Research Updates",
                   "5" = "Administrating Vaccines",
                   "6" = "Student Vaccination",
                   "7" = "National Covid Death Reports",
                   "8" = "Vaccine Approval"))
  
  dominant_topic_2021 <- gamma_2021_k10 %>%
    group_by(document) %>%
    slice_max(gamma, n = 1, with_ties = FALSE) %>%
    ungroup()
  
  topic_counts_2021 <- dominant_topic_2021 %>%
    count(topic, sort = TRUE) %>% 
    mutate(topic = reorder(topic, n))
  
  ggplot(topic_counts_2021, aes(x = topic, y = n, fill = topic)) +
    geom_bar(stat = "identity", width = 1, color = "white", show.legend = F) +
    geom_text(aes(label = n)) +
    coord_flip() +
    labs(
      title = "The Biggest Themes of April - June 2021",
      subtitle = "Relating to Covid-19",
      fill = "Themes",
      x = NULL,
      y = "Tweets"
    ) + theme_bw()
  
```
In contrast to the previous topics found in tweets from mid-2020, where the prevalent themes regarded most of official covid updates, such as infection counts, new policy and big changes, and health guidelines as well as government responses, the topics of 2021 is dominated by the existence of vaccines. Obviously, there was no covid vaccine while investigating the tweets from 2020, since the first covid vaccine came out in the December of 2020. 

**Some context:** At this time, we were under the Biden administration, which joins COVAX---program aimed at vaccinating people in low and middle income countries against COVID-19. Hence the second most prevalent topic in 2021 regarding covid: global vaccine rollouts. Poorer countries did not have access to vaccines as early as we did. While experimenting with other k values, I saw 'India' pop up a lot.

There is strong pushback for vaccines in general, but during 2021, variants like *delta* and *omicron* came to be. They were either extremely more contagious or more severe, but a large population of people still refused to wear masks and get vaccinated, fearing the long term complications due to limited testing. Some other, more niche topics I discovered and found interesting while reading some tweets to help gauge the context of some words during the beta stage (but I changed k-values and lost their focuses) were...
1. research was being done to find covid's origins
2. a lawsuit towards a Texas hospital for requiring covid vaccinations
3. mask mandates and pushback

Ultimately, comparing the tweets and their topics with a one year gap highlight vaccine developments. In the 2020 tweets, Florida was just starting to reopen. And in 2021, life felt more normal, at least for the U.S. (hence less conversations on the national death reports.) The same cannot be said about lower income countries globally, which relied of other countries' distribution of vaccines. Despite vaccines being a game changer, allowing the reopening of society to function so well, masks were and vaccines still are very controversial. The theme of vaccines heavily dominated several topics in 2021.

## Conclusion

Jadyn.
